### 2025-04-20 code/analysis-code/analysis.qmd
Part 5 analysis

---
title: "Analysis"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(survey)
library(tableone)
library(mice)
library(mitools)
library(miceadds)
library(tidymodels)
library(caret)
library(gt)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
#data <- readRDS(here("data","processed-data", "data3.rds"))
#data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems

data_imputed <- readRDS(here("data","processed-data", "binary_imputed_datasets.rds"))
```

As a reminder, we are going to use the following variables, and stratify by covid test result.
EXERCISE: Whether or not individual met guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met, 1 if any were met
SLEEP: If individual reported getting the proper amount of sleep for their resspective age group
DEPRESSED: If individual felt depressed weekly+ or not
TROUBLE_SLEEPING: If individual reported having trouble sleeping over past several days, more than half days, or nearly every day or not at all
SOCIAL: If person felt like they were reciveing less social support or not
COVID: If they've tested positive for covid or not

We create estimates and CIs using complex survery design variables, using the multiple imputed data.

```{r}

# Create survey design objects for each transformed dataset
svy_designs <- lapply(data_imputed, function(data) {
  svydesign(ids = ~PSU, strata = ~STRATA, weights = ~SAMPWEIGHT, data = data , nest=TRUE)
})

# Function to get complete estimates (both 0 and 1) stratified by COVID status
get_complete_stratified_estimates <- function(var_name) {
  results <- data.frame(
    Variable = character(),
    COVID_Status = character(),
    Binary_Value = numeric(),
    Estimate = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    stringsAsFactors = FALSE
  )

  # Loop through COVID status (0 and 1)
  for (covid_val in c(0, 1)) {
    covid_label <- ifelse(covid_val == 0, "Negative", "Positive")

    # Calculate proportions for this COVID status
    props <- lapply(svy_designs, function(design) {
      # Create a subset design
      subdesign <- subset(design, COVID == covid_val)
      # Create formula for the variable
      formula <- as.formula(paste0("~factor(", var_name, ")"))
      # Calculate proportions
      svymean(formula, subdesign, na.rm = TRUE)
    })

    # Pool results
    pooled <- MIcombine(props)

    # Extract results for both binary values (if available)
    coef_names <- names(coef(pooled))

    # Handle different potential naming patterns
    for (bin_val in c(0, 1)) {
      # Check different possible coefficient names
      possible_names <- c(
        paste0("factor(", var_name, ")", bin_val),
        as.character(bin_val)
      )

      # Find the matching coefficient name
      coef_idx <- NULL
      for (name in possible_names) {
        if (name %in% coef_names) {
          coef_idx <- which(coef_names == name)
          break
        }
      }

      # If found, extract estimates
      if (!is.null(coef_idx)) {
        est <- coef(pooled)[coef_idx]
        ci <- confint(pooled)[coef_idx, ]

        # Add to results
        results <- rbind(results, data.frame(
          Variable = var_name,
          COVID_Status = covid_label,
          Binary_Value = bin_val,
          Estimate = round(est * 100, 1),
          CI_lower = round(ci[1] * 100, 1),
          CI_upper = round(ci[2] * 100, 1)
        ))
      }
    }
  }

  return(results)
}

# Variables to analyze
variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Get complete stratified estimates for each variable
complete_results <- lapply(variables, get_complete_stratified_estimates)

# Combine results into a single data frame
complete_df <- do.call(rbind, complete_results)

# Reshape for better presentation
formatted_results <- complete_df %>%
  mutate(Binary_Label = ifelse(Binary_Value == 1, "Yes", "No")) %>%
  select(Variable, COVID_Status, Binary_Label, Estimate, CI_lower, CI_upper) %>%
  arrange(Variable, COVID_Status, desc(Binary_Label))

print(formatted_results)

```

Now let's turn that into a nice table using GT, and I'll save it to use in the manuscript later. 
```{r}
#checking
summary(formatted_results)
str(formatted_results)
names(formatted_results)

# Prepare the data by combining Variable and Binary_Label, pivoting, and formatting the results
table1_imputed <- formatted_results %>%
  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),  # Convert Yes to 1 and No to 0
         Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
  pivot_wider(
    names_from = COVID_Status,
    values_from = c(Estimate, CI_lower, CI_upper),
    names_glue = "{COVID_Status}_{.value}"
  ) %>%
  mutate(
    # Scale the estimates and CI by 0.01
    Estimate_Negative = Negative_Estimate * 0.01,
    CI_Negative = sprintf("(%.2f, %.2f)", Negative_CI_lower * 0.01, Negative_CI_upper * 0.01),
    Estimate_Positive = Positive_Estimate * 0.01,
    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01)
  ) %>%
  select(Variable_Label, Estimate_Negative, CI_Negative, Estimate_Positive, CI_Positive) %>%
  gt() %>%
  tab_header(
    title = "Estimated proportions of mental health related factors by COVID test status"
  ) %>%
  tab_spanner(
    label = "COVID Negative",
    columns = c(Estimate_Negative, CI_Negative)
  ) %>%
  tab_spanner(
    label = "COVID Positive",
    columns = c(Estimate_Positive, CI_Positive)
  ) %>%
  cols_label(
    Variable_Label = "Variable",               # Rename Variable_Label to Variable
    Estimate_Negative = "Estimate",            # Rename Estimate_Negative to Estimate
    CI_Negative = "95% CI",                    # Rename CI_Negative to 95% CI
    Estimate_Positive = "Estimate",            # Rename Estimate_Positive to Estimate
    CI_Positive = "95% CI"                     # Rename CI_Positive to 95% CI
  )

# Save the table as a PNG
here()
gtsave(table1_imputed, here("results", "figures", "table1_imputed.png"))
```
################################################################################
Now, we will start modeling COVID postivity as an outcome of mental health factors.

Up first, we will create singleton models to individually study the effects of mental health factors on covid test positivity status.

```{r}
# Create a survey design object for each completed dataset
survey_designs <- lapply(data_imputed, function(data_i) {
  svydesign(
    id = ~PSU,
    strata = ~STRATA,
    weights = ~SAMPWEIGHT,
    nest = TRUE,
    data = data_i
  )
})

# Fit svyglm models for EXERCISE predictor
models_exercise <- lapply(survey_designs, function(dsgn) {
  svyglm(COVID ~ EXERCISE, design = dsgn, family = quasibinomial())
})

# Extract coefficients and standard errors for each model
model_exercise_coef <- lapply(models_exercise, function(model) {
  data.frame(
    Estimate = coef(model),
    Std.Error = sqrt(diag(vcov(model))),
    stringsAsFactors = FALSE
  )
})

# Calculate means and variances for pooling
estimates <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Estimate))
se <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Std.Error))

# Compute the mean across imputations
mean_estimate <- apply(estimates, 2, mean)

# Compute within-imputation variance (average variance across imputations)
within_variance <- apply(se^2, 2, mean)

# Compute between-imputation variance (variance of the estimates across imputations)
between_variance <- apply(estimates, 2, var)

# Calculate the total variance using Rubin's rule
total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance

# Calculate the pooled standard error
pooled_se <- sqrt(total_variance)

# Calculate the 95% confidence intervals
ci_low <- mean_estimate - 1.96 * pooled_se
ci_high <- mean_estimate + 1.96 * pooled_se

# Create a data frame with pooled estimates, standard errors, and confidence intervals
pooled_exercise <- data.frame(
  Estimate = mean_estimate,
  Std.Error = pooled_se,
  CI_Low = ci_low,
  CI_High = ci_high
)

# Display the pooled results
pooled_exercise

library(survey)

# Fit svyglm models for EXERCISE predictor (and other predictors if needed)
predictors <- c("EXERCISE", "DEPRESSED", "SLEEP", "TROUBLE_SLEEPING", "SOCIAL")

# Fit models for each predictor
models <- lapply(predictors, function(predictor) {
  lapply(survey_designs, function(dsgn) {
    formula <- as.formula(paste("COVID ~", predictor))
    svyglm(formula, design = dsgn, family = quasibinomial())
  })
})

# Extract coefficients and standard errors for each model
model_coef <- mapply(function(predictor, predictor_models) {
  lapply(predictor_models, function(model) {
    data.frame(
      Estimate = coef(model),
      Std.Error = sqrt(diag(vcov(model))),
      stringsAsFactors = FALSE
    )
  })
}, predictors, models, SIMPLIFY = FALSE)

# Calculate means, variances, and p-values for pooling
pooled_results <- lapply(names(model_coef), function(predictor_name) {
  coefs <- model_coef[[predictor_name]]

  estimates <- do.call(rbind, lapply(coefs, function(x) x$Estimate))
  se <- do.call(rbind, lapply(coefs, function(x) x$Std.Error))

  # Compute the mean across imputations
  mean_estimate <- apply(estimates, 2, mean)

  # Compute within-imputation variance (average variance across imputations)
  within_variance <- apply(se^2, 2, mean)

  # Compute between-imputation variance (variance of the estimates across imputations)
  between_variance <- apply(estimates, 2, var)

  # Calculate the total variance using Rubin's rule
  total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance

  # Calculate the pooled standard error
  pooled_se <- sqrt(total_variance)

  # Calculate the 95% confidence intervals
  ci_low <- mean_estimate - 1.96 * pooled_se
  ci_high <- mean_estimate + 1.96 * pooled_se

  # Calculate z-statistics and p-values
  z_stat <- mean_estimate / pooled_se
  p_values <- 2 * (1 - pnorm(abs(z_stat)))  # Two-tailed p-value

  # Exponentiate to get Odds Ratios (ORs) from log-odds
  or_estimate <- exp(mean_estimate)
  ci_low_or <- exp(ci_low)
  ci_high_or <- exp(ci_high)

  # Create a data frame with pooled Odds Ratios, standard errors, confidence intervals, and p-values
  data.frame(
    Predictor = rep(predictor_name, length(mean_estimate)),
    OR = or_estimate,
    Std.Error = pooled_se,
    CI_Low = ci_low_or,
    CI_High = ci_high_or,
    p_value = p_values,
    row.names = names(mean_estimate)
  )
})

# Combine results for all predictors into one data frame
pooled_exercise <- do.call(rbind, pooled_results)

# Display the labeled pooled results
pooled_exercise

```

Now we will make a nice table of that using gt.

```{r}
names(pooled_exercise)
str(pooled_exercise)

#not including intercepts values
# Assuming 'pooled_exercise' is your data frame
pooled_exercise %>%
  group_by(Predictor) %>%
  slice(2) %>%  # Select only the second row for each Predictor
  ungroup() %>%
  mutate(CI = paste0("(", round(CI_Low, 3), ", ", round(CI_High, 3), ")")) %>%
  select(Predictor, OR, CI, p_value) %>%
  gt() %>%
  tab_header(
    title = "Associations of individual mental health factors with positive covid test result"
  ) %>%
  cols_label(
    Predictor = "Predictor",
    OR = "Odds Ratio",
    CI = "Confidence Interval",
    p_value = "P-Value"
  ) %>%
  fmt_number(
    columns = vars(OR, p_value),
    decimals = 3
  )

```

################################################################################
This is the part of the project in which I demonstrate complex techniques learned during this class.

Next, we will make the multivariate model. I will use cross validation for this. The RMSE is found to be 0.4450602.
```{r}

cross_validation_imputed_rmse <- function(imputed_data, design, predictors, k_folds = 5) {
  rmse_list <- numeric(length(imputed_data))  # Store RMSE for each imputation

  for (i in 1:length(imputed_data)) {
    data <- imputed_data[[i]]
    set.seed(123)
    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)

    fold_rmse <- numeric(k_folds)  # RMSE for each fold

    for (fold in 1:k_folds) {
      train_data <- data[folds != fold, ]
      test_data  <- data[folds == fold, ]

      # Survey design for training and testing
      train_design <- svydesign(ids = ~1, data = train_data, weights = ~SAMPWEIGHT)
      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
      model <- svyglm(formula, design = train_design, family = quasibinomial())

      # Predictions
      predictions <- predict(model, newdata = test_data, type = "response")
      actual <- test_data$COVID

      # Compute RMSE for this fold
      fold_rmse[fold] <- sqrt(mean((predictions - actual)^2, na.rm = TRUE))
    }

    # Average RMSE over folds for this imputed dataset
    rmse_list[i] <- mean(fold_rmse)
  }

  # Average RMSE over all imputed datasets
  overall_rmse <- mean(rmse_list)

  return(list(rmse_per_imputation = rmse_list, overall_rmse = overall_rmse))
}

# Assuming you have your imputed_data and a vector of predictors
results <- cross_validation_imputed_rmse(imputed_data = data_imputed,
                                         design = NULL,  # Not used inside, can be omitted
                                         predictors = c("AGE", "SLEEP", "DEPRESSED"))

# View results
results$rmse_per_imputation  # RMSE for each imputed dataset
results$overall_rmse         # Average RMSE across imputations
```

Now, we will make a LASSO regression model. For this, I'm not going to use the multiple imputation dataset. I'm just going to pick a sinlge imputed dataset, the first one. The RMSE is 0.276.

```{r}
# Extract the first completed imputed dataset
data1 <- data_imputed[[1]]  # You can loop over 1:5 later

# Then define the survey design
survey_design <- svydesign(
  ids = ~PSU,
  strata = ~STRATA,
  weights = ~SAMPWEIGHT,
  data = data1,
  nest=TRUE
)

set.seed(1234)  # for reproducibility

# Choose your outcome and predictors
outcome_var <- "DEPRESSED"
predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Loop over imputations and compute RMSE
rmse_list <- map(data_imputed, function(dat) {
  # Drop missing in predictors or outcome if any slipped through
  dat <- dat %>% drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))

  # Create design matrix X and response y
  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
  y <- dat[[outcome_var]]
  w <- dat$SAMPWEIGHT

  # Split into training and test sets (e.g., 70/30 split)
  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
  test_idx <- setdiff(seq_len(nrow(X)), train_idx)

  X_train <- X[train_idx, ]
  y_train <- y[train_idx]
  w_train <- w[train_idx]

  X_test <- X[test_idx, ]
  y_test <- y[test_idx]
  w_test <- w[test_idx]

  # Fit weighted lasso model
  cv_fit <- cv.glmnet(
    x = X_train,
    y = y_train,
    alpha = 1,
    weights = w_train,
    family = "binomial",  # change to "binomial" for binary outcome
    nfolds = 5
  )

  # Predict on test set
  preds <- predict(cv_fit, newx = X_test, s = "lambda.min")

  # Weighted RMSE
  sqrt(weighted.mean((y_test - preds)^2, w_test))
})

# Pool RMSE across imputations
pooled_rmse <- mean(unlist(rmse_list))

# Output
cat("RMSE for each imputation:\n")
print(unlist(rmse_list))
cat("\nPooled RMSE across imputations:", round(pooled_rmse, 3), "\n")

```

################################################################################
The following code was used prior to development of imputed data and is no longer relevant. It may be ignored. I am keeping it in case it serves as a useful reference for the rest of the project or future endeavors.
################################################################################

We create estimates and CIs using complex survey design variables
```{r}
#list variables
varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

#list categories
cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

survey = svydesign(
  ids = ~PSU, 
  strata = ~STRATA, 
  weights = ~SAMPWEIGHT, 
  nest = TRUE, 
  data = data)

exercise <- svyby(~as.factor(EXERCISE), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
sleep <- svyby(~as.factor(SLEEP), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
depressed <- svyby(~as.factor(DEPRESSED), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
trouble_sleeping <- svyby(~as.factor(TROUBLE_SLEEPING), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
social <- svyby(~as.factor(SOCIAL), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
str(exercise)
```

Let's rearrange these exercise numbers into a table. I used ChatGPT and Claude to help with making the table. 
```{r}
# Convert the svyby object to a data frame (if not already)
exercise_df <- as.data.frame(exercise)

# Clean up column names to make them more readable
colnames(exercise_df) <- c("COVID", "EXERCISE_0", "EXERCISE_1", "CI_L_EXERCISE_0", "CI_L_EXERCISE_1", "CI_U_EXERCISE_0", "CI_U_EXERCISE_1")

# Use a different approach to reshape the data
exercise_long <- exercise_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and exercise level
  mutate(
    Type = case_when(
      grepl("^EXERCISE", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Exercise = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Exercise),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
exercise_with_ci <- exercise_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Exercise, Mean, CI)

# Now create the final table format
exercise_final <- exercise_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Exercise,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
exercise_table <- exercise_final %>%
  gt() %>%
  tab_header(
    title = "Exercise Data by COVID Status"
  ) %>%
  cols_label(
    Exercise = "Exercise Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Exercise == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise Level: 0 = No regular exercise, 1 = Meets requirements for regular exercise according to American Heart Association",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Exercise),
    fn = function(x) {
      ifelse(x == "0", "No Regular Exercise", "Regular Exercise")
    }
  )

# Print the table
exercise_table
```

Next, we will add the sleep variable. I used ChatGPT for this.
```{r}
# Convert the svyby object to a data frame (if not already)
sleep_df <- as.data.frame(sleep)  # Assuming 'sleep' is the dataset

# Clean up column names to make them more readable
colnames(sleep_df) <- c("COVID", "SLEEP_0", "SLEEP_1", "CI_L_SLEEP_0", "CI_L_SLEEP_1", "CI_U_SLEEP_0", "CI_U_SLEEP_1")

# Use a different approach to reshape the data
sleep_long <- sleep_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and sleep level
  mutate(
    Type = case_when(
      grepl("^SLEEP", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Sleep = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Sleep),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
sleep_with_ci <- sleep_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Sleep, Mean, CI)

# Now create the final table format
sleep_final <- sleep_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Sleep,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
sleep_table <- sleep_final %>%
  gt() %>%
  tab_header(
    title = "Sleep Data by COVID Status"
  ) %>%
  cols_label(
    Sleep = "Sleep Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Sleep == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Sleep Level: 0 = Less than recommended sleep, 1 = Met recommendations for sleep by age groups according to NIH",
    locations = cells_column_labels(columns = Sleep)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Sleep),
    fn = function(x) {
      ifelse(x == "0", "Not enough sleep", "Enough sleep")
    }
  )

# Print the table
sleep_table

#################################################################################

# Extract the data from the gt tables
exercise_table_df <- as.data.frame(exercise_table)
sleep_table_df <- as.data.frame(sleep_table)

# Ensure that the column names are the same across both tables
colnames(sleep_table_df) <- colnames(exercise_table_df)

# Combine the tables row by row
combined_df <- bind_rows(exercise_table_df, sleep_table_df)

# Convert the combined data frame back to a gt table
combined_table <- combined_df %>%
  gt() %>%
  tab_header(
    title = "Combined Exercise and Sleep Data by COVID Status"
  )

# Print the combined gt table
combined_table
```

Create the same tables with depressed, trouble_sleeping, and social. This was completed in BBEdit.
```{r}
#########################################################################################
# depressed variable

# Convert the svyby object to a data frame (if not already)
depressed_df <- as.data.frame(depressed)

# Clean up column names to make them more readable
colnames(depressed_df) <- c("COVID", "depressed_0", "depressed_1", "CI_L_depressed_0", "CI_L_depressed_1", "CI_U_depressed_0", "CI_U_depressed_1")

# Use a different approach to reshape the data
depressed_long <- depressed_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and depressed level
  mutate(
    Type = case_when(
      grepl("^depressed", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    depressed = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, depressed),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
depressed_with_ci <- depressed_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, depressed, Mean, CI)

# Now create the final table format
depressed_final <- depressed_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = depressed,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
depressed_table <- depressed_final %>%
  gt() %>%
  tab_header(
    title = "depressed Data by COVID Status"
  ) %>%
  cols_label(
    depressed = "depressed Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = depressed == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Depressed level: 0 = Feeling depressed less than weekly, 1 = Feeling depressed weekly or more",
    locations = cells_column_labels(columns = depressed)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = depressed),
    fn = function(x) {
      ifelse(x == "0", "Not regularly depressed", "Frequent depressed feeling")
    }
  )

# Print the table
depressed_table

#########################################################################################
# trouble_sleeping variable

# Convert the svyby object to a data frame (if not already)
trouble_sleeping_df <- as.data.frame(trouble_sleeping)

# Clean up column names to make them more readable
colnames(trouble_sleeping_df) <- c("COVID", "trouble_sleeping_0", "trouble_sleeping_1", "CI_L_trouble_sleeping_0", "CI_L_trouble_sleeping_1", "CI_U_trouble_sleeping_0", "CI_U_trouble_sleeping_1")

# Use a different approach to reshape the data
trouble_sleeping_long <- trouble_sleeping_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and trouble_sleeping level
  mutate(
    Type = case_when(
      grepl("^trouble_sleeping", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    trouble_sleeping = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, trouble_sleeping),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
trouble_sleeping_with_ci <- trouble_sleeping_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, trouble_sleeping, Mean, CI)

# Now create the final table format
trouble_sleeping_final <- trouble_sleeping_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = trouble_sleeping,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
trouble_sleeping_table <- trouble_sleeping_final %>%
  gt() %>%
  tab_header(
    title = "trouble_sleeping Data by COVID Status"
  ) %>%
  cols_label(
    trouble_sleeping = "trouble_sleeping Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = trouble_sleeping == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "trouble_sleeping level: 0 = No trouble sleeping, 1 = Reported trouble sleeping",
    locations = cells_column_labels(columns = trouble_sleeping)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = trouble_sleeping),
    fn = function(x) {
      ifelse(x == "0", "No reported trouble sleeping", "Regular trouble sleeping")
    }
  )

# Print the table
trouble_sleeping_table

#########################################################################################
# social variable

# Convert the svyby object to a data frame (if not already)
social_df <- as.data.frame(social)

# Clean up column names to make them more readable
colnames(social_df) <- c("COVID", "social_0", "social_1", "CI_L_social_0", "CI_L_social_1", "CI_U_social_0", "CI_U_social_1")

# Use a different approach to reshape the data
social_long <- social_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and social level
  mutate(
    Type = case_when(
      grepl("^social", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    social = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, social),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
social_with_ci <- social_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, social, Mean, CI)

# Now create the final table format
social_final <- social_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = social,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
social_table <- social_final %>%
  gt() %>%
  tab_header(
    title = "social Data by COVID Status"
  ) %>%
  cols_label(
    social = "social Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = social == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "social Level: 0 = Regular social support, 1 = Less social support",
    locations = cells_column_labels(columns = social)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = social),
    fn = function(x) {
      ifelse(x == "0", "Regular social support", "Less social support")
    }
  )

# Print the table
social_table

```

Now, we will create the combined table of exercise, sleep, depressed, trouble_sleeping, and social. ChatGPT helped with many tweaks.
```{r}
# Extract the data from the other gt tables
depressed_table_df <- as.data.frame(depressed_table)
trouble_sleeping_table_df <- as.data.frame(trouble_sleeping_table)
social_table_df <- as.data.frame(social_table)

# Ensure the column names match across all tables
colnames(depressed_table_df) <- colnames(combined_df)
colnames(trouble_sleeping_table_df) <- colnames(combined_df)
colnames(social_table_df) <- colnames(combined_df)

# Combine all the tables row by row
combined_all_df <- bind_rows(combined_df, depressed_table_df, trouble_sleeping_table_df, social_table_df)

# Add a row_id column to the data frame
combined_all_df <- combined_all_df %>%
  mutate(row_id = row_number())

# Convert the combined data frame back to a gt table
combined_all_table <- combined_all_df %>%
  gt() %>%
  tab_header(
    title = "Mental Health Factors by SARS-CoV-2 Test Status"
  ) %>%
  # Spanners for COVID Negative and COVID Positive
  tab_spanner(
    label = "Never has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  tab_spanner(
    label = "Has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  # Remove the word 'Exercise' from the first row
  cols_label(
    Exercise = ""  # This removes the "Exercise" label
  ) %>%
  # Format the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group.",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add alternating row colors using row_id
  tab_style(
    style = cell_fill(color = "#f5f5f5"),  # Gray color for odd rows
    locations = cells_body(rows = row_id %% 2 == 1)  # Apply to odd rows
  ) %>%
  tab_style(
    style = cell_fill(color = "white"),  # White color for even rows
    locations = cells_body(rows = row_id %% 2 == 0)  # Apply to even rows
  ) %>%
  # Add solid vertical line between row names and Mean_0 columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_0")
    )
  ) %>%
  # Add solid vertical line between COVID Negative and COVID Positive columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_1")
    )
  ) %>%
  # Hide the row_id column
  cols_hide(columns = "row_id")

# Print the combined gt table
combined_all_table

#footnote:
#"Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group."

# Save the table as a PNG
here()
gtsave(combined_all_table, here("results", "figures", "result_table1.png"))
```

################################################################################

Up next, we will run the singleton models. For these, our outcome variable (covid test result) will be used as a variable for each of the respective predictor vairbales that we created in the processing code. These models appear as the following:

$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Sleep} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Depressed} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{TroubleSleeping} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Social} $$
Create survey design.
```{r}
#review names to use below
names(data)

#create survey object using design variables
survey_design <- svydesign(
  ids = ~PSU,
  strata = ~STRATA,
  weights = ~SAMPWEIGHT,
  data = data,
  nest = TRUE
)

pr_results <- data.frame()

variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
               "SOCIAL")

for (var in variables) {
  # Create the formula dynamically
  formula <- as.formula(paste("COVID ~", var))

  # Fit the model
  model <- svyglm(
    formula, 
    design = survey_design, 
    family = quasibinomial(link = "log")
  )

  # Extract prevalence ratios and confidence intervals
  pr <- exp(coef(model))  # Prevalence ratios
  pr_ci <- exp(confint(model))  # Confidence intervals

  # Extract p-values
  p_values <- summary(model)$coefficients[, "Pr(>|t|)"]

  # Combine results into a single table
  pr_results <- rbind(
    pr_results,
    data.frame(
      Variable = var,
      Estimate = pr[-1],  # Skip the intercept
      `2.5% CI` = pr_ci[-1, 1],
      `97.5% CI` = pr_ci[-1, 2],
      `p-value` = p_values[-1]  # Skip the intercept
    )
  )
}

# Print the results
print(pr_results)

```

Now, we will make that into a nice quality table using the gt package and the help of chatGPT and Claude.
```{r}
#checking
glimpse(pr_results)
str(pr_results)

#create the gt table
pr_results %>%
  mutate(
    Predictor = stringr::str_to_title(gsub("_", " ", Variable)),
    `Prevalence Ratio` = round(Estimate, 2),
    `95% CI` = paste0("(", round(X2.5..CI, 2), ", ", round(X97.5..CI, 2), ")"),
    `p-value` = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(Predictor, `Prevalence Ratio`, `95% CI`, `p-value`) %>%
  gt() %>%
  tab_header(
    title = md("**Positive COVID Test Result Association with Mental Health Factors**")
  ) %>%
  tab_options(
    column_labels.border.bottom.width = px(1.5),  # Bold line under column headers
    column_labels.border.bottom.color = "black"
  )

```

Now, we're going to complete the process of model selection, using backwards selection. We want to end up with a model we can use for prediction purposes. Our initial model looks like this: 
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} + \beta_2 \cdot X_{Sleep} + \beta_3 \cdot X_{Depressed} + \beta_4 \cdot X_{TroubleSleeping} + \beta_5 \cdot X_{Social} $$
Let's find the AIC value. 
```{r}

unique(data$COVID)
model_full <- svyglm(
  COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL,
  design = survey_design,
  family = quasibinomial(link = "log")
)
# Perform backward selection based on AIC
model_step <- step(model_full, direction = "backward", trace = 1)
summary(model_step)

```