### 2025-04-30 products/manuscript/Manuscript.qmd

---
title: "Mental Health Factors and Covid"
author: "Annalise Cramer"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
#bibliography: ../../assets/dataanalysis-references.bib
#csl: ../../assets/american-journal-of-epidemiology.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
library(mice)
library(miceadds)
library(mitools)
library(tableone)
library(survey)
library(gt)
here()
```

**Authors**

* Annalise Cramer$^{1,2 \land}$ (ORCID: 0009-0002-8718-3593)

**Author affiliations**

1. College of Public Health, University of Georgia, Athens, GA, USA.
2. Center for Ecology of Infectious Diseases, Athens, GA, USA.

$\land$ Corresponding author: atc10260@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect those of the University of Georgia.

{{< pagebreak >}}

# Summary/Abstract

Isolation and quarantine strategies employed for containment of SARS-Cov-2 infection led to unexpected mental health struggles among the population of the United States. Infected individuals and healthy individuals were affected, but little is known about the potential relationships between mental health and physical health. In this study, we utilized Nataional Health Information Surveys from 2019-2021 to examine the relationships between receieving a positive COVID-19 test and self reporting mental health factors such as meeting recommended guidelines for exercise, meeting the recommended number of hours of sleep per night, reporting feeling depressed recently, reporting having trouble sleeping recently, and reporting receiving less social support.

We find

{{< pagebreak >}}

# Introduction

## General Background Information

When the SARS-CoV-2 pandemic began in late 2019, there were very few control measures enacted within the United States. By Spring of 2020, this had drastically changed, and control measures were broadly put into effect across the country. Social distancing, quantining, isolation, and lockdowns were prioritized control of transmission, but societal impacts were not initially considered. Disruptions to daily routine and social interactions quickly became a widespread concerns for their potential impacts on mental health, with widespread increases in anxiety, depression, loneliness, and sleep disturbances (CITATION NEEDED).

Pandemic control strategies have become commonly criticized for the impacts on individuals' psychological well-being. Research documents the mental health consequences of pandemic control strageies (CITATION SAME AS BEFORE), but the potential relationships between mental health factors and vulnerability to SARS-CoV-2 infection is less understood.

This study seeks to add to the limited understanding of mental health and COVID-19 vulnerability by expliorung associations between mental health factors and COVID-19 test positivity status. Self reported bouts of depressed mood, perceived social support, amount of exercise, and sleep quality (via hours of sleep and reporting difficulty falling asleep) are examined. (CITATION HERE) Simple and multiple regression models including Lasso and Random Forest are used to explore the predicitve power of mental health factors to a positive COVID-19 test.

Cite this properly later:
1.) https://www.tandfonline.com/doi/full/10.1080/09638237.2020.1757052

## Description of data and data source

Data for this project is survey data collected from the National Health Interview Survey, produced by IPUMS through the University of Minnesota. This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States.

## Questions/Hypotheses to be addressed

1.) What is the prevalence of poor mental health?
2.) Is poor mental health associated with a positive covid test?

Factors studied include:
1.) Average self-reported hours of sleep per night, 
2.) Self-reported depressed mood, 
3.) Self-reported trouble sleeping, 
4.) Self-reported average weekly exercise, and
 5.) Self-reported changes to social and emotional support.

{{< pagebreak >}}

# Methods

## Schematic of workflow

```{r}
#| label: fig-schematic
#| fig-cap: "Flowchart depicting methodology"
#| echo: FALSE
knitr::include_graphics(here("results","figures", "flowchart.png"))
```

## Data aquisition

National Health Information Survey data was obtained from IPUMS, used with permission. Data was subsetted to years 2019-2022, to cover the brunt of pandemic years. Despite low prevelance of COVID-19 cases in 2019 worldwide, the appearance of COVID-19 in the news (and consequential potential impact on mental health factors) led us to include this data. The final dataset contains 151406 observations.

## Data import and cleaning

Data was used in 2020 and after, given the NHIS didn't have questions about COVID-19 testing in 2019, these questions began in 2020. Binary variables were created to allow for interpretability.

EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whether weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy. For adults, this includes 150 minutes per week of moderate intensity aerobic activity, 75 minutes per week of vigorous aerobic activity, or a combination of both.Children ages 6-17 should get 60 minutes of moderate to vigorous intenisty physical activity. A value of 1 for this binary variable means that the individual met exercise recommendations, and a 0 means that the individual did not meet exercise recommendations.

SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age. Individuals under age 1 are suggested to get 12-16 hours of sleep per day, children ages 1 to 2 are suggested to get 11-14 hours of sleep per day, children ages 3 to 5 years are suggested to get 10-13 hours of sleep per day, children ages 6 to 12 years are suggested to get 9-12 hours of sleep per day, teens ages 13-18 are suggested to get 8-10 hours of sleep per day, and adults (over the age of 18) are suggested to get 7-9 hours of sleep per day. Guidelines do not require that sleep takes place at nighttime, and details that naps count towards sleep totals. A value of 1 for this binary variable means that the individual met recommendations for hours of sleep per day, and a 0 means that the individual did not meet recommendations for hours of sleep recommendations.

DEPRESSED includes a 1 if the individual reports feeling depressed on a weekly or daily basis, but a 0 for feeling depressed at a frequency of monthly or less.

TROUBLE_SLEEPING includes a 1 if the individual reports struggling to fall asleep several days per week, more than half the days, or nearly every day, and 0 if they report otherwise.

SOCIAL includes a value of 1 for an individual describing feeling like they have less social support than their regular amount, and a 0 for feeling like they recieve the same or more support.

For all variables, answers for don't know or unsure were converted to missing.

## Statistical analysis

After creating binary variables, significant amounts of data were missing, seen in table below. The proportions of values for answers 1 and 0 for binary variables are shown out out total answers, and then a column including NAs are shown, which details the proportions of NA answers among total individuals included.

Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis. To calculate later statistics, processes were preformed on each of the 5 sets and then averaged.

```{r}
#| label: tab-1
#| fig-cap: "Missingness table"
#| echo: FALSE
knitr::include_graphics(here("results","figures", "summary_table_with_nas.png"))
```

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT. A summary table was generated to find occurence of each of the factors and their inverses among the data. Next, single models were ran to find associations between each variable with covid test postivity. To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.

```{r}
#| label: fig-1
#| fig-cap: "Summary statistics table"
#| echo: FALSE
knitr::include_graphics(here("results","figures", "table1_imputed.png"))
```

## Basic statistical analysis

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Single model table"
#| echo: FALSE
here()
knitr::include_graphics(here("results","figures", "single_model_table.png"))
```

## Full analysis

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

{{< pagebreak >}}

# References