### 2025-05-01 code/analysis-code/analysis.qmd
Part 6 updates

---
title: "Analysis"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(survey)
library(tableone)
library(mice)
library(mitools)
library(miceadds)
library(tidymodels)
library(caret)
library(gt)
library(scales)
library(pROC)
library(ranger)
library(glmnet)
library(Metrics)
library(purrr)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
#data <- readRDS(here("data","processed-data", "data3.rds"))
#data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems

data_imputed <- readRDS(here("data","processed-data", "binary_imputed_datasets.rds"))
```

As a reminder, we are going to use the following variables, and stratify by covid test result.
EXERCISE: Whether or not individual met guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met, 1 if any were met
SLEEP: If individual reported getting the proper amount of sleep for their resspective age group
DEPRESSED: If individual felt depressed weekly+ or not
TROUBLE_SLEEPING: If individual reported having trouble sleeping over past several days, more than half days, or nearly every day or not at all
SOCIAL: If person felt like they were reciveing less social support or not
COVID: If they've tested positive for covid or not

We create estimates and CIs using complex survery design variables, using the multiple imputed data.

```{r}

# Create survey design objects for each transformed dataset
svy_designs <- lapply(data_imputed, function(data) {
  svydesign(ids = ~PSU, strata = ~STRATA, weights = ~SAMPWEIGHT, data = data , nest=TRUE)
})

# Function to get complete estimates (both 0 and 1) stratified by COVID status
get_complete_stratified_estimates <- function(var_name) {
  results <- data.frame(
    Variable = character(),
    COVID_Status = character(),
    Binary_Value = numeric(),
    Estimate = numeric(),
    CI_lower = numeric(),
    CI_upper = numeric(),
    stringsAsFactors = FALSE
  )

  # Loop through COVID status (0 and 1)
  for (covid_val in c(0, 1)) {
    covid_label <- ifelse(covid_val == 0, "Negative", "Positive")

    # Calculate proportions for this COVID status
    props <- lapply(svy_designs, function(design) {
      # Create a subset design
      subdesign <- subset(design, COVID == covid_val)
      # Create formula for the variable
      formula <- as.formula(paste0("~factor(", var_name, ")"))
      # Calculate proportions
      svymean(formula, subdesign, na.rm = TRUE)
    })

    # Pool results
    pooled <- MIcombine(props)

    # Extract results for both binary values (if available)
    coef_names <- names(coef(pooled))

    # Handle different potential naming patterns
    for (bin_val in c(0, 1)) {
      # Check different possible coefficient names
      possible_names <- c(
        paste0("factor(", var_name, ")", bin_val),
        as.character(bin_val)
      )

      # Find the matching coefficient name
      coef_idx <- NULL
      for (name in possible_names) {
        if (name %in% coef_names) {
          coef_idx <- which(coef_names == name)
          break
        }
      }

      # If found, extract estimates
      if (!is.null(coef_idx)) {
        est <- coef(pooled)[coef_idx]
        ci <- confint(pooled)[coef_idx, ]

        # Add to results
        results <- rbind(results, data.frame(
          Variable = var_name,
          COVID_Status = covid_label,
          Binary_Value = bin_val,
          Estimate = round(est * 100, 1),
          CI_lower = round(ci[1] * 100, 1),
          CI_upper = round(ci[2] * 100, 1)
        ))
      }
    }
  }

  return(results)
}

# Variables to analyze
variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Get complete stratified estimates for each variable
complete_results <- lapply(variables, get_complete_stratified_estimates)

# Combine results into a single data frame
complete_df <- do.call(rbind, complete_results)

# Reshape for better presentation
formatted_results <- complete_df %>%
  mutate(Binary_Label = ifelse(Binary_Value == 1, "Yes", "No")) %>%
  select(Variable, COVID_Status, Binary_Label, Estimate, CI_lower, CI_upper) %>%
  arrange(Variable, COVID_Status, desc(Binary_Label))

print(formatted_results)

```

Now let's turn that into a nice table using GT, and I'll save it to use in the manuscript later. 
```{r}
#checking
summary(formatted_results)
str(formatted_results)
names(formatted_results)

# Prepare the data by combining Variable and Binary_Label, pivoting, and formatting the results
table1_imputed <- formatted_results %>%
  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),  # Convert Yes to 1 and No to 0
         Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
  pivot_wider(
    names_from = COVID_Status,
    values_from = c(Estimate, CI_lower, CI_upper),
    names_glue = "{COVID_Status}_{.value}"
  ) %>%
  mutate(
    # Scale the estimates and CI by 0.01
    Estimate_Negative = Negative_Estimate * 0.01,
    CI_Negative = sprintf("(%.2f, %.2f)", Negative_CI_lower * 0.01, Negative_CI_upper * 0.01),
    Estimate_Positive = Positive_Estimate * 0.01,
    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01)
  ) %>%
  select(Variable_Label, Estimate_Negative, CI_Negative, Estimate_Positive, CI_Positive) %>%
  gt() %>%
  tab_header(
    title = "Estimated proportions of mental health related factors by COVID test status"
  ) %>%
  tab_spanner(
    label = "COVID Negative",
    columns = c(Estimate_Negative, CI_Negative)
  ) %>%
  tab_spanner(
    label = "COVID Positive",
    columns = c(Estimate_Positive, CI_Positive)
  ) %>%
  cols_label(
    Variable_Label = "Variable",               # Rename Variable_Label to Variable
    Estimate_Negative = "Estimate",            # Rename Estimate_Negative to Estimate
    CI_Negative = "95% CI",                    # Rename CI_Negative to 95% CI
    Estimate_Positive = "Estimate",            # Rename Estimate_Positive to Estimate
    CI_Positive = "95% CI"                     # Rename CI_Positive to 95% CI
  )

# Save the table as a PNG
here()
gtsave(table1_imputed, here("results", "figures", "table1_imputed.png"))
```
################################################################################
Now, we will start modeling COVID postivity as an outcome of mental health factors.

Up first, we will create singleton models to individually study the effects of mental health factors on covid test positivity status.

```{r}
# Create a survey design object for each completed dataset
survey_designs <- lapply(data_imputed, function(data_i) {
  svydesign(
    id = ~PSU,
    strata = ~STRATA,
    weights = ~SAMPWEIGHT,
    nest = TRUE,
    data = data_i
  )
})

# Fit svyglm models for EXERCISE predictor (and other predictors if needed)
predictors <- c("EXERCISE", "DEPRESSED", "SLEEP", "TROUBLE_SLEEPING", "SOCIAL")

# Fit models for each predictor
models <- lapply(predictors, function(predictor) {
  lapply(survey_designs, function(dsgn) {
    formula <- as.formula(paste("COVID ~", predictor))
    svyglm(formula, design = dsgn, family = quasibinomial())
  })
})

# Extract coefficients and standard errors for each model
model_coef <- mapply(function(predictor, predictor_models) {
  lapply(predictor_models, function(model) {
    data.frame(
      Estimate = coef(model),
      Std.Error = sqrt(diag(vcov(model))),
      stringsAsFactors = FALSE
    )
  })
}, predictors, models, SIMPLIFY = FALSE)

# Calculate means, variances, and p-values for pooling
pooled_results <- lapply(names(model_coef), function(predictor_name) {
  coefs <- model_coef[[predictor_name]]

  estimates <- do.call(rbind, lapply(coefs, function(x) x$Estimate))
  se <- do.call(rbind, lapply(coefs, function(x) x$Std.Error))

  # Compute the mean across imputations
  mean_estimate <- apply(estimates, 2, mean)

  # Compute within-imputation variance (average variance across imputations)
  within_variance <- apply(se^2, 2, mean)

  # Compute between-imputation variance (variance of the estimates across imputations)
  between_variance <- apply(estimates, 2, var)

  # Calculate the total variance using Rubin's rule
  total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance

  # Calculate the pooled standard error
  pooled_se <- sqrt(total_variance)

  # Calculate the 95% confidence intervals
  ci_low <- mean_estimate - 1.96 * pooled_se
  ci_high <- mean_estimate + 1.96 * pooled_se

  # Calculate z-statistics and p-values
  z_stat <- mean_estimate / pooled_se
  p_values <- 2 * (1 - pnorm(abs(z_stat)))  # Two-tailed p-value

  # Exponentiate to get Odds Ratios (ORs) from log-odds
  or_estimate <- exp(mean_estimate)
  ci_low_or <- exp(ci_low)
  ci_high_or <- exp(ci_high)

  # Create a data frame with pooled Odds Ratios, standard errors, confidence intervals, and p-values
  data.frame(
    Predictor = rep(predictor_name, length(mean_estimate)),
    OR = or_estimate,
    Std.Error = pooled_se,
    CI_Low = ci_low_or,
    CI_High = ci_high_or,
    p_value = p_values,
    row.names = names(mean_estimate)
  )
})

# Combine results for all predictors into one data frame
pooled_exercise <- do.call(rbind, pooled_results)

# Display the labeled pooled results
pooled_exercise

```

Now we will make a nice table of that using gt and save it to use in the mauscript.

```{r}
names(pooled_exercise)
str(pooled_exercise)

#not including intercepts values
table2 <- pooled_exercise %>%
  group_by(Predictor) %>%
  slice(2) %>%  # Select only the second row for each Predictor
  ungroup() %>%
  mutate(CI = paste0("(", round(CI_Low, 3), ", ", round(CI_High, 3), ")")) %>%
  select(Predictor, OR, CI, p_value) %>%
  gt() %>%
  tab_header(
    title = "Associations of individual mental health factors with positive covid test result"
  ) %>%
  cols_label(
    Predictor = "Predictor",
    OR = "Odds Ratio",
    CI = "Confidence Interval",
    p_value = "P-Value"
  ) %>%
  fmt_number(
    columns = vars(OR, p_value),
    decimals = 3
  )

# Save the table as a PNG
here()
gtsave(table2, here("results", "figures", "single_model_table.png"))

```

################################################################################
This is the part of the project in which I demonstrate complex techniques learned during this class.

Next, we will make the multivariate model. I will use cross validation for this.Pooled AUC is calculated and a observed-expected plot is generated.
```{r}
library(pROC)
library(dplyr)

cross_validation_imputed_auc_logit <- function(imputed_data, predictors, k_folds = 5) {
  auc_list <- numeric(length(imputed_data))  # Store AUC for each imputation

  for (i in seq_along(imputed_data)) {
    data <- imputed_data[[i]]

    # Drop rows with missing values in outcome, predictors, or weights
    data <- data %>% drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))

    set.seed(123)
    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)

    fold_auc <- numeric(k_folds)

    for (fold in 1:k_folds) {
      train_data <- data[folds != fold, ]
      test_data  <- data[folds == fold, ]

      # Fit logistic regression model
      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
      logit_model <- glm(
        formula = formula,
        data = train_data,
        family = binomial(),
        weights = SAMPWEIGHT
      )

      # Predict probabilities on test data
      pred_probs <- predict(logit_model, newdata = test_data, type = "response")
      actual <- test_data$COVID

      # Compute AUC for this fold
      auc_result <- roc(actual, pred_probs)$auc
      fold_auc[fold] <- auc_result
    }

    auc_list[i] <- mean(fold_auc)
  }

  overall_auc <- mean(auc_list)

  return(list(auc_per_imputation = auc_list, overall_auc = overall_auc))
}

# Usage
results_logit <- cross_validation_imputed_auc_logit(
  imputed_data = data_imputed,
  predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
)

# View results
results_logit$auc_per_imputation
results_logit$overall_auc

# Function to compute calibration data per imputation
get_calibration_data <- function(data, predictors, n_bins = 10) {
  # Fit logistic regression model
  formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
  model <- glm(formula, data = data, family = binomial(), weights = SAMPWEIGHT)

  # Predict probabilities
  data$predicted <- predict(model, type = "response")

  # Bin predictions into deciles
  data <- data %>%
    mutate(bin = ntile(predicted, n_bins)) %>%
    group_by(bin) %>%
    summarise(
      mean_pred = mean(predicted),
      obs_rate = mean(COVID),
      .groups = "drop"
    )

  return(data)
}

# Combine calibration data across imputations
all_calib_data <- purrr::map_dfr(
  .x = seq_along(data_imputed),
  .f = function(i) {
    data <- data_imputed[[i]] %>% 
      drop_na(COVID, EXERCISE, SLEEP, DEPRESSED, TROUBLE_SLEEPING, SOCIAL, SAMPWEIGHT)

    calib <- get_calibration_data(data, predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL"))
    calib$imputation <- paste0("Imputation ", i)
    return(calib)
  }
)

# Plot calibration curves
LogReg_plot <- ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
  geom_line() +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  labs(title = "Calibration Plot: Logistic Regression (All 5 Imputations)", color = "Imputation") +
  theme_minimal()

# Save the plot using the `here` package
ggsave(filename = here::here("results", "figures", "log_calibration_plot.png"),
       plot = LogReg_plot,
       width = 8, height = 6, dpi = 300)

```

Now, we will make a LASSO regression model. For this, I'm going to use the multiple imputation dataset with cross validation with 5 folds. Instead of using complex survey design, I will weight the model by sampleweight. Pooled AUC is calculated and a observed-expected plot is generated.

```{r}
library(glmnet)
library(caret)
library(Metrics)  # for rmse()
library(dplyr)

set.seed(1234)
k <- 5
auc_list <- c()

# Loop through 5 imputed datasets
for (i in 1:5) {
  data1 <- data_imputed[[i]]

  # Remove rows with missing outcome
  data1_complete <- data1 %>% filter(!is.na(COVID))

  # Create folds
  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)

  preds <- rep(NA, nrow(data1_complete))

  for (fold_idx in seq_along(folds)) {
    test_idx <- folds[[fold_idx]]
    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)

    train_data <- data1_complete[train_idx, ] %>% drop_na()
    test_data  <- data1_complete[test_idx, ] %>% drop_na()

    # Create model matrix (remove outcome column)
    x_train <- model.matrix(COVID ~ ., data = train_data)[, -1]
    y_train <- train_data$COVID

    x_test <- model.matrix(COVID ~ ., data = test_data)[, -1]

    # Fit Lasso with CV
    lasso_fit <- cv.glmnet(x = x_train, y = y_train, alpha = 1, family = "binomial")

    # Predict probabilities (type = "response" for prob)
    preds[test_idx] <- predict(lasso_fit, newx = x_test, s = "lambda.min", type = "response")
  }

  actuals <- data1_complete$COVID

  # AUC using pROC
  roc_obj <- roc(actuals, preds)
  fold_auc <- auc(roc_obj)
  auc_list[i] <- fold_auc
}

pooled_auc <- mean(auc_list)
print(pooled_auc)

################################################################################
## plot expected vs observed

set.seed(1234)
k <- 5
all_preds_df <- data.frame()

for (i in 1:5) {
  data1 <- data_imputed[[i]]
  data1_complete <- data1 %>% filter(!is.na(COVID))

  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)

  preds <- rep(NA, nrow(data1_complete))

  for (fold_idx in seq_along(folds)) {
    test_idx <- folds[[fold_idx]]
    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)

    train_data <- data1_complete[train_idx, ] %>% drop_na()
    test_data  <- data1_complete[test_idx, ] %>% drop_na()

    x_train <- model.matrix(COVID ~ ., data = train_data)[, -1]
    y_train <- train_data$COVID
    x_test  <- model.matrix(COVID ~ ., data = test_data)[, -1]

    lasso_fit <- cv.glmnet(x = x_train, y = y_train, alpha = 1, family = "binomial")

    preds[test_idx] <- predict(lasso_fit, newx = x_test, s = "lambda.min", type = "response")
  }

  all_preds_df <- bind_rows(all_preds_df, data.frame(
    imputation = i,
    predicted = preds,
    observed = data1_complete$COVID
  ))
}

# Bin predictions into 10 bins and compute calibration stats
calib_data <- all_preds_df %>%
  mutate(bin = ntile(predicted, 10)) %>%
  group_by(imputation, bin) %>%
  summarise(
    mean_pred = mean(predicted, na.rm = TRUE),
    obs_rate = mean(observed, na.rm = TRUE),
    .groups = "drop"
  )

Lasso_plot <- ggplot(calib_data, aes(x = mean_pred, y = obs_rate, color = factor(imputation))) +
  geom_line() +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  labs(
    title = "Calibration Plot: Lasso Logistic Regression (All 5 Imputations)",
    color = "Imputation"
  ) +
  theme_minimal()

print(Lasso_plot)

# Save the plot using the `here` package
ggsave(filename = here::here("results", "figures", "lasso_calibration_plot.png"),
       plot = Lasso_plot,
       width = 8, height = 6, dpi = 300)

################################################################################
# Keep variable names for predictors
formula <- as.formula("COVID ~ .")

# Explicitly remove rows with any NA in predictors (even if COVID is not NA)
train_data <- train_data %>% select(COVID, everything()) %>% drop_na()

# Now get model matrix
x_train <- model.matrix(formula, data = train_data)[, -1]
y_train <- train_data$COVID

nrow(x_train)
length(y_train)
```

```{r}
# Load required libraries
library(glmnet)      # For Lasso regression
library(pROC)        # For AUC calculation
library(ggplot2)     # For plotting
library(dplyr)       # For data manipulation
library(mice)        # For working with imputed data
library(tidyr)       # For data reshaping

# --- LASSO MODEL WITH CROSS-VALIDATION ON IMPUTED DATA ---

# Function to run Lasso with sample weights and calculate performance metrics
run_lasso_cv <- function(imp_data, outcome = "COVID", 
                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
                        weight_var = "sampleweight") {

  # Extract data from current imputation
  X <- as.matrix(imp_data[, predictors])
  y <- imp_data[[outcome]]
  weights <- imp_data[[weight_var]]

  # Standardize predictors (important for Lasso)
  X_scaled <- scale(X)

  # Run cross-validated Lasso with sample weights
  set.seed(123) # For reproducibility
  cv_model <- cv.glmnet(
    x = X_scaled,
    y = y,
    family = "binomial",
    alpha = 1,       # 1 = Lasso
    weights = weights,
    nfolds = 5,
    type.measure = "auc"
  )

  # Get the best lambda (penalty parameter)
  best_lambda <- cv_model$lambda.min

  # Refit model with best lambda
  final_model <- glmnet(
    x = X_scaled,
    y = y,
    family = "binomial",
    alpha = 1,
    weights = weights,
    lambda = best_lambda
  )

  # Get predictions
  pred_probs <- predict(final_model, newx = X_scaled, type = "response")

  # Calculate AUC
  roc_obj <- roc(y, pred_probs)
  auc_value <- auc(roc_obj)

  # Create calibration data for expected vs observed plot
  n_groups <- 10  # Number of bins for calibration plot

  # Create calibration data
  calib_data <- data.frame(
    predicted = as.vector(pred_probs),
    observed = y
  )

  # Sort by predicted probability and divide into bins
  calib_data <- calib_data[order(calib_data$predicted), ]
  calib_data$bin <- cut(seq(1, nrow(calib_data)), breaks = n_groups, labels = FALSE)

  # Calculate mean predicted and observed values for each bin
  calibration <- calib_data %>%
    group_by(bin) %>%
    summarize(
      n = n(),
      mean_predicted = mean(predicted),
      mean_observed = mean(observed)
    )

  # Return results
  return(list(
    model = final_model,
    lambda = best_lambda,
    auc = auc_value,
    calibration = calibration,
    coefs = coef(final_model)
  ))
}

# ----------------------------
# MAIN ANALYSIS
# ----------------------------

# Check the structure of the imputed data
print(class(data_imputed))

# Initialize results containers
auc_values <- numeric(length = 5)
calibration_results <- list()
coef_results <- list()

# Function to run Lasso with sample weights and calculate performance metrics
run_lasso_cv <- function(imp_data, outcome = "COVID", 
                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
                        weight_var = "sampleweight") {

  # Extract data from current imputation and handle missing values
  # First, check for and handle missing data
  complete_cases <- complete.cases(imp_data[, c(outcome, predictors, weight_var)])
  if (!all(complete_cases)) {
    message("Removing ", sum(!complete_cases), " rows with missing values")
    imp_data <- imp_data[complete_cases, ]
  }

  X <- as.matrix(imp_data[, predictors])
  y <- imp_data[[outcome]]
  weights <- imp_data[[weight_var]]

  # Standardize predictors (important for Lasso)
  X_scaled <- scale(X)

  # Run cross-validated Lasso with sample weights
  set.seed(123) # For reproducibility
  cv_model <- cv.glmnet(
    x = X_scaled,
    y = y,
    family = "binomial",
    alpha = 1,       # 1 = Lasso
    weights = weights,
    nfolds = 5,
    type.measure = "auc"
  )

  # Get the best lambda (penalty parameter)
  best_lambda <- cv_model$lambda.min

  # Refit model with best lambda
  final_model <- glmnet(
    x = X_scaled,
    y = y,
    family = "binomial",
    alpha = 1,
    weights = weights,
    lambda = best_lambda
  )

# Combine calibration results from all imputations
all_calibration <- do.call(rbind, calibration_results)

# ----------------------------
# OUTPUT RESULTS
# ----------------------------

# Print AUC values for each imputation
cat("\nAUC values for each imputation:\n")
print(auc_values)
cat("\nMean AUC across imputations:", mean(auc_values), "\n")

# Plot expected vs observed probabilities for all imputations
calibration_plot <- ggplot(all_calibration, aes(x = mean_predicted, y = mean_observed, color = factor(imputation))) +
  geom_point(size = 3) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  labs(
    title = "Expected vs. Observed Probabilities",
    subtitle = paste("Mean AUC =", round(mean(auc_values), 3)),
    x = "Expected (Predicted Probability)",
    y = "Observed (Actual Proportion)",
    color = "Imputation"
  ) +
  theme_minimal() +
  coord_equal() +
  xlim(0, 1) + 
  ylim(0, 1)

print(calibration_plot)

# Print average coefficients across imputations
cat("\nAverage Lasso coefficients across imputations:\n")
# Extract coefficient names from the first imputation
coef_names <- rownames(coef_results[[1]])

# Create matrix to store coefficients from all imputations
all_coefs <- matrix(0, nrow = length(coef_names), ncol = 5)
rownames(all_coefs) <- coef_names

# Fill in the coefficient matrix
for (i in 1:5) {
  all_coefs[, i] <- as.vector(coef_results[[i]])
}

# Calculate mean coefficients
mean_coefs <- rowMeans(all_coefs)
coef_summary <- data.frame(
  Variable = coef_names,
  Mean_Coefficient = mean_coefs
)

# Print non-zero coefficients
non_zero_coefs <- coef_summary[abs(coef_summary$Mean_Coefficient) > 0.001, ]
print(non_zero_coefs[order(abs(non_zero_coefs$Mean_Coefficient), decreasing = TRUE), ])
```

```{r}
library(dplyr)
library(ggplot2)
library(scales)

calib_data_list <- vector("list", length(data_imputed))

for (i in seq_along(data_imputed)) {
  dat <- data_imputed[[i]] %>%
    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))

  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
  y <- dat[[outcome_var]]
  w <- dat$SAMPWEIGHT

  set.seed(1234 + i)
  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
  test_idx <- setdiff(seq_len(nrow(X)), train_idx)

  X_train <- X[train_idx, ]
  y_train <- y[train_idx]
  w_train <- w[train_idx]

  X_test <- X[test_idx, ]
  y_test <- y[test_idx]

  cv_fit <- cv.glmnet(
    x = X_train,
    y = y_train,
    weights = w_train,
    alpha = 1,
    family = "binomial",
    standardize = TRUE
  )

  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()

  # Count unique predicted probabilities
  n_unique_preds <- length(unique(pred_probs))

  # Only proceed with binning if there are at least 2 unique predicted probabilities
  if (n_unique_preds >= 2) {
    # Safely bin the predicted probabilities
    n_bins <- min(5, n_unique_preds)

    calib_data <- tibble(
      actual = y_test,
      predicted = pred_probs
    ) %>%
      mutate(bin = cut_number(predicted, n = n_bins, labels = FALSE)) %>%
      group_by(bin) %>%
      summarise(
        mean_pred = mean(predicted),
        obs_rate = mean(actual),
        .groups = "drop"
      ) %>%
      mutate(imputation = paste0("Imputation ", i))

    calib_data_list[[i]] <- calib_data
  } else {
    message("Skipping calibration plot for imputation ", i, ": too few unique predicted values (", n_unique_preds, " unique values)")
  }
}

# Combine all calibration data
all_calib_data <- bind_rows(calib_data_list)

# Create calibration plot
ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
  geom_line() +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
  theme_minimal()

# Combine calibration data
all_calib_data <- bind_rows(calib_data_list)

# Plot calibration curve
ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
  geom_line() +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
  theme_minimal()

# Combine calibration data
all_calib_data <- bind_rows(calib_data_list)

# Plot calibration curve
ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
  geom_line() +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
  theme_minimal()
# Save the plot
ggsave(filename = here::here("results", "figures", "lasso_calibration_plot.png"),
       plot = calibration_plot_lasso,
       width = 8, height = 6, dpi = 300)

# Initialize list to store AUCs
auc_list <- vector("numeric", length(data_imputed))

for (i in seq_along(data_imputed)) {
  dat <- data_imputed[[i]] %>%
    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))

  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
  y <- dat[[outcome_var]]
  w <- dat$SAMPWEIGHT

  set.seed(1234 + i)
  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
  test_idx <- setdiff(seq_len(X), train_idx)

  X_train <- X[train_idx, ]
  y_train <- y[train_idx]
  w_train <- w[train_idx]

  X_test <- X[test_idx, ]
  y_test <- y[test_idx]

  # Fit lasso model
  cv_fit <- cv.glmnet(
    x = X_train,
    y = y_train,
    weights = w_train,
    alpha = 1,
    family = "binomial",
    standardize = TRUE
  )

  # Predict probabilities
  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()

  # Compute AUC
  auc_val <- roc(y_test, pred_probs)$auc
  auc_list[i] <- auc_val
}

# Print AUCs
cat("AUC for each imputation:\n")
print(round(auc_list, 3))

# Mean (pooled) AUC
cat("\nPooled AUC across imputations:", round(mean(auc_list), 3), "\n")
```

Finally, we will construct a random forest model. For this, I will not use the multiple imputation dataset, as before, I will use the first imputed dataset, with cross validation 5 folds. Pooled AUC is calculated and a observed-expected plot is generated.

```{r}
cross_validation_imputed_predictions <- function(imputed_data, predictors, k_folds = 5) {
  auc_list <- numeric(length(imputed_data))  # Store AUCs
  predictions_list <- vector("list", length(imputed_data))  # Store predictions and actuals

  for (i in 1:length(imputed_data)) {
    data <- imputed_data[[i]] %>%
      drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))

    set.seed(123)
    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)

    fold_auc <- numeric(k_folds)
    imputation_preds <- tibble()  # Empty tibble to collect preds for this imputation

    for (fold in 1:k_folds) {
      train_data <- data[folds != fold, ]
      test_data  <- data[folds == fold, ]

      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
      rf_model <- ranger(
        formula = formula,
        data = train_data,
        case.weights = train_data$SAMPWEIGHT,
        probability = TRUE
      )

      pred_probs <- predict(rf_model, data = test_data)$predictions[, 2]
      actual <- test_data$COVID

      # Store predictions
      imputation_preds <- bind_rows(imputation_preds, tibble(actual = actual, predicted = pred_probs))

      # AUC
      auc_result <- roc(actual, pred_probs)$auc
      fold_auc[fold] <- auc_result
    }

    auc_list[i] <- mean(fold_auc)
    predictions_list[[i]] <- imputation_preds
  }

  list(
    auc_per_imputation = auc_list,
    overall_auc = mean(auc_list),
    predictions = predictions_list  # List of tibbles with actual + predicted
  )
}
# Assuming you have your imputed_data and a vector of predictors
results_with_preds <- cross_validation_imputed_predictions(
  imputed_data = data_imputed,
  predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
)

# View results
results$auc_per_imputation  # AUC for each imputed dataset
results$overall_auc         # Average AUC across imputations

# Create calibration plot from predictions across imputations
plot_calibration_multiple <- function(predictions_list, n_bins = 10, title = "Calibration Plot") {
  calib_data_all <- map2_dfr(predictions_list, seq_along(predictions_list), function(df, i) {
    df %>%
      mutate(bin = ntile(predicted, n_bins)) %>%
      group_by(bin) %>%
      summarise(
        mean_pred = mean(predicted),
        obs_rate = mean(actual),
        .groups = "drop"
      ) %>%
      mutate(imputation = paste0("Imputation ", i))
  })

  ggplot(calib_data_all, aes(x = mean_pred, y = obs_rate, color = imputation)) +
    geom_line() +
    geom_point(size = 2) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
    scale_x_continuous("Expected probability", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    scale_y_continuous("Observed proportion", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    labs(title = title, color = "Imputation") +
    theme_minimal()
}

# Plot it
plot_cal <- plot_calibration_multiple(results_with_preds$predictions, title = "Calibration Plot: RF across 5 Imputations")
print(plot_cal)

# Save it
ggsave(here("results", "figures", "rf_calibration_all_imputations.png"), plot_cal, width = 8, height = 6, dpi = 300)

```

Let's put these into a nice table for the manuscript. 
```{r}
# Create a data frame with the results
auc_table <- data.frame(
  Model = c("Multivariate", "Lasso", "Random Forest"),
  AUC = c(0.523, 0.529, 0.556)
)

# Create and style the GT table
model_table <- auc_table %>%
  gt() %>%
  tab_header(
    title = "Model Performance Comparison",
  ) %>%
  fmt_number(
    columns = vars(AUC),
    decimals = 3
  )

# Print the table
model_table

# Save the table as a PNG
here()
gtsave(model_table, here("results", "figures", "model_table.png"))

# Store your AUC lists here
aauc_multivar <- results$auc_per_imputation  # From multivariate logistic regression with CV
auc_lasso     <- unlist(auc_list)                      # From your LASSO section
auc_rf        <- auc_list                              # From your final RF model section

# Create a data frame with AUC values and model type
auc_df <- data.frame(
  Imputation = rep(1:5, 3),
  AUC = c(auc_multivar, auc_lasso, auc_rf),
  Model = rep(c("Multivariate Logistic Regression", "LASSO", "Random Forest"), each = 5)
)

# Plot: AUCs across imputations by model
aucs <- ggplot(auc_df, aes(x = Model, y = AUC, color = Model)) +
  geom_jitter(width = 0.2, size = 2) +         # dots for individual AUCs
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, color = "black") +
  stat_summary(fun = mean, geom = "text", aes(label = round(..y.., 3)), 
               vjust = -1, color = "black", fontface = "bold") +
  labs(title = "AUC Across Imputations by Model",
       x = "Model",
       y = "AUC") +
  theme_minimal()

# Save the table as a PNG
here()
ggsave(
  filename = here("results", "figures", "auc_plot.png"),
  plot = aucs,  # replace with your plot object name
  width = 8,
  height = 6,
  dpi = 300
)

```

Now, we will make expected versus observed plots.
```{r}
plot_calibration_multiple <- function(actual, predicted_list, n_bins = 10, title = "Calibration Plot") {
  # Combine calibration data across imputations
  calib_data_all <- map2_dfr(predicted_list, seq_along(predicted_list), function(predicted, i) {
    tibble(actual = actual, predicted = predicted) %>%
      mutate(bin = ntile(predicted, n_bins)) %>%
      group_by(bin) %>%
      summarise(
        mean_pred = mean(predicted),
        obs_rate = mean(actual),
        .groups = "drop"
      ) %>%
      mutate(imputation = paste0("Imputation ", i))
  })

  # Plot
  ggplot(calib_data_all, aes(x = mean_pred, y = obs_rate, color = imputation)) +
    geom_line() +
    geom_point(size = 2) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
    scale_x_continuous("Expected probability", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    scale_y_continuous("Observed proportion", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    labs(title = title, color = "Imputation") +
    theme_minimal()
}

# Example usage:
plot_cal <- plot_calibration_multiple(
  actual = test_data$COVID,
  predicted_list = pred_probs_list,  # this should be a list of 5 vectors
  title = "Random Forest Calibration Across 5 Imputations"
)

# Display plot
print(plot_cal)

# Save it
ggsave(
  filename = here("results", "figures", "rf_calibration_all_imputations.png"),
  plot = plot_cal,
  width = 8,
  height = 6,
  dpi = 300
)

```